Multi-target detection with application to cryo-electron microscopy

### 2019-02-15 inverse_problems/inverse_problems.tex
no message

Let $Z\in \R^m$ be a random signal drawn from some fixed distribution.  The autocorrelation of order $q = 1, 2, \ldots$ is given for any integer shifts $\ell_1, \ldots, \ell_{q-1}$ by
\begin{equation}
a_Z^q[\ell_1,\ldots,\ell_{q-1}]   = \E\left\{\frac{1}{m} \sum_{i=-\infty}^{\infty} Z[i]Z[i+\ell_1]\cdots Z[i+\ell_{q-1}]\right\},
\label{eq:ac_general}
\end{equation}
where the expectation is taken with respect to the distribution of $Z$. The indexing of $Z$ out of the range $0, \ldots, m-1$ is zero-padded.
Explicitly, the first-, second- and third-order autocorrelations are given by: 
\begin{align} 
a_Z^1 & = \frac{1}{m} \sum_{i=0}^{m-1} \E\left\{
Z[i]\right\}, \nonumber\\
a_Z^2[\ell] & = \frac{1}{m} \sum_{i = \max\{0, -\ell\}}^{m-1 + \min\{0, -\ell\}} \E\left\{Z[i]Z[i+\ell]\right\},\\
a_Z^3[\ell_1,\ell_2] & = \frac{1}{m} \sum_{i = \max\{0, -\ell_1, -\ell_2\}}^{m-1 + \min\{0, -\ell_1, -\ell_2\}} \E\left\{Z[i]Z[i+\ell_1]Z[i+\ell_2]\right\}.  \nonumber \label{eq:ac_special}
\end{align}
The autocorrelation functions have symmetries. Specifically, $$\E\left\{a_Z^2[\ell]\right\} = \E\left\{a_Z^2[-\ell]\right\},$$ and \TODO{to verify}
$$\E\left\{a_Z^3[\ell_1,\ell_2]\right\} = \E\left\{a_Z^3[\ell_2,\ell_1]\right\}=\E\left\{a_Z^3[-\ell_1,\ell_2-\ell_1]\right\}.
$$
For our purposes, this will be applied both to $x$ (of length $L$) and to $y$ (of length $N$).

@@

Along the same lines, one can establish a relation for third-order autocorrelations:
\begin{equation}
 \label{eq:ac3_micrograph}
\lim_{N\to\infty} a_y^3[\ell_1,\ell_2] \aseq \gamma a_{x}^3[\ell_1,\ell_2]  + \sigma^2\gamma a_{X}^1  \big(\delta[\ell_1,0]+\delta[0,\ell_2]
+\delta[\ell_1,\ell_2]\big),
\end{equation}
for $\ell_1,\ell_2 = 0, \ldots, L-1$, where $\delta[\ell_1,\ell_2]:=\delta[\ell_1-\ell_2]$.
Here too, few entries are affected by $\sigma$ in the limit.

@@

Autocorrelations in $d$ dimensions is defined for  $\ell_1,\ldots,\ell_{q-1}\in\mathbb{Z}^d$  as 
\begin{equation}
a_Z^q[\ell_1,\ldots,\ell_{q-1}]   = \E\left\{\frac{1}{m^d} \sum_{i\in\mathbb{Z}^d} z[i]z[i+\ell_1]\cdots z[i+\ell_{q-1}]\right\}.
\label{eq:ac_d_dimension}
\end{equation}
Interestingly, for dimensions greater than one almost all deterministic (that is, the distribution of $Z$ is a Delta function) signals are determined uniquely from their second-order autocorrelation, up to two symmetries: sign (or phase for complex signals) and reflection through the origin (with conjugation in the complex case)~\cite{hayes1982reconstruction}. 
If the mean of signal is available and non-zero, the sign symmetry can be resolved. However, determining the reflection symmetry still requires additional information, beyond the second-order autocorrelation.
The case of 1-D signals is essentially different: generally there are $2^{L-2}$ signals with the same second-order autocorrelation (after eliminating  symmetries)~\cite{beinert2015ambiguities,bendory2017fourier}.

>>>>

In what follows, we consider autocorrelations of both the observation $y$ and of the signal occurrences in $y$. As per our discussion of extensions, the signal occurrences may be sampled from a discrete set $\{x_1, \ldots, x_K\}$, or from a more general distribution. Accordingly, we define autocorrelations for a random signal $z$ of length $m$ \TODO{change $m$ to something upper case, maybe $U$}. For our purposes, this will be applied both to signal occurrences (of length $L$) and to $y$ (of length $N$).

For a random signal $z\in \R^m$, the autocorrelation of order $q = 1, 2, \ldots$ is given for any integer shifts $\ell_1, \ldots, \ell_{q-1}$ by
\begin{equation}
a_z^q[\ell_1,\ldots,\ell_{q-1}]   = \frac{1}{m} \sum_{i=-\infty}^{\infty} \E\left\{z[i]z[i+\ell_1]\cdots z[i+\ell_{q-1}]\right\},
\label{eq:ac_general}
\end{equation}
where the expectation is taken with respect to the distribution of $z$. Indexing of $z$ out of the range $0, \ldots, m-1$ is zero-padded.
Explicitly, the first-, second- and third-order autocorrelations are given by: 
\begin{align} 
a_z^1 & = \frac{1}{m} \sum_{i=0}^{m-1} \E\left\{
z[i]\right\}, \nonumber\\
a_z^2[\ell] & = \frac{1}{m} \sum_{i = \max\{0, -\ell\}}^{m-1 + \min\{0, -\ell\}} \E\left\{z[i]z[i+\ell]\right\},\\
a_z^3[\ell_1,\ell_2] & = \frac{1}{m} \sum_{i = \max\{0, -\ell_1, -\ell_2\}}^{m-1 + \min\{0, -\ell_1, -\ell_2\}} \E\left\{z[i]z[i+\ell_1]z[i+\ell_2]\right\}.  \nonumber \label{eq:ac_special}
\end{align}
Since  autocorrelations depend only on the differences between indices, the obey the following symmetries: $$a_z^2[\ell] = a_z^2[-\ell],$$ and
$$a_z^3[\ell_1,\ell_2] = a_z^3[\ell_2,\ell_1]=a_z^3[-\ell_1,\ell_2-\ell_1].
$$

@@

Along the same lines, one can establish a relation for third-order autocorrelations:
\begin{equation}
 \label{eq:ac3_micrograph}
\lim_{N\to\infty} a_y^3[\ell_1,\ell_2] \aseq \gamma a_{X}^3[\ell_1,\ell_2]  + \sigma^2\gamma a_{X}^1  \big(\delta[\ell_1,0]+\delta[0,\ell_2]
+\delta[\ell_1,\ell_2]\big),
\end{equation}
for $\ell_1,\ell_2 = 0, \ldots, L-1$, where $\delta[\ell_1,\ell_2]:=\delta[\ell_1-\ell_2]$.
Here too, few entries are affected by $\sigma$ in the limit.

In particular, for $X$ sampled from $\{ x_1, \ldots, x_K \}$ with probabilities $(\pi_1, \ldots, \pi_K)$, the autocorrelations are given in explicit form as:
\begin{align} 
a_X^1 & = \frac{1}{m} \sum_{i=0}^{m-1} \E\left\{
X[i]\right\}, \nonumber\\
a_X^2[\ell] & = \frac{1}{m} \sum_{i = \max\{0, -\ell\}}^{m-1 + \min\{0, -\ell\}} \E\left\{X[i]X[i+\ell]\right\},\\
a_X^3[\ell_1,\ell_2] & = \frac{1}{m} \sum_{i = \max\{0, -\ell_1, -\ell_2\}}^{m-1 + \min\{0, -\ell_1, -\ell_2\}} \E\left\{X[i]X[i+\ell_1]X[i+\ell_2]\right\}.  \nonumber
\end{align}
\TODO{ABOVE: replace expectations with sums.}

@@

Autocorrelations in $d$ dimensions is defined for  $\ell_1,\ldots,\ell_{q-1}\in\mathbb{Z}^d$  as 
\begin{equation}
a_z^q[\ell_1,\ldots,\ell_{q-1}]   = \E\left\{\frac{1}{m^d} \sum_{i\in\mathbb{Z}^d} z[i]z[i+\ell_1]\cdots z[i+\ell_{q-1}]\right\}.
\label{eq:ac_d_dimension}
\end{equation}
Interestingly, for dimensions greater than one almost all deterministic (that is, the distribution of $z$ is a Delta function) signals are determined uniquely from their second-order autocorrelation, up to two symmetries: sign (or phase for complex signals) and reflection through the origin (with conjugation in the complex case)~\cite{hayes1982reconstruction}. 
If the mean of signal is available and non-zero, the sign symmetry can be resolved. However, determining the reflection symmetry still requires additional information, beyond the second-order autocorrelation.
The case of 1-D signals is essentially different: generally there are $2^{L-2}$ signals with the same second-order autocorrelation (after eliminating  symmetries)~\cite{beinert2015ambiguities,bendory2017fourier}.
